---
title: "Factorial ANOVA"
author: "Gery Casiez"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(apa)
```

## Factorial ANOVA using 2 factors

Between-subject design 
Factor 1: text entry method (standard word-processing software, text entry using word-prediction software, text entry using speech-based dictation software)
Factor 2: task (transcription, composition)

## Load data

```{r readdata}
data = read.table("wordProcessingSoftwareBetweenSubjectFactorialANOVA2factors.csv", header=TRUE, sep=",")
kable(data)
```


Convert data using the long format
```{r convertlong, results="hide", message=FALSE}
library("reshape")
data.long = melt(data, id = c("participant","textmethod","task","time"))
```

Indicate what are the independent variables (factors)
```{r indepvar, results="hide"}
data.long$textmethod = factor(data.long$textmethod)
data.long$task = factor(data.long$task)
```

## ANOVA
```{r anova}
library("ez")
kable(anova_apa(ezANOVA(data.long, dv=.(time), wid=.(participant), between=.(textmethod,task), detailed=TRUE)))
```

## Plot


```{r plottask}
library(dplyr)
library(gmodels)
library(ggplot2)

plot_data = data %>%  group_by(task) %>% summarise(meanTime = mean(time), ci.lower = ci(time)["CI lower"], ci.upper = ci(time)["CI upper"])
ggplot(plot_data, aes(x = factor(task), y = meanTime)) +
      geom_bar(stat="identity") +
      geom_errorbar(aes(ymin = ci.lower, ymax = ci.upper), width = .1, position = position_dodge(.05)) +
      theme(panel.background = element_blank(),
            panel.border = element_rect(colour = 'black', fill=NA),
            panel.grid.major.y = element_line( size = .1, color = "grey")
            ) +
      labs(title = "", x = "Task", y = "Time")

```

```{r plot}

plot_data = data %>%  group_by(textmethod) %>% summarise(meanTime = mean(time), ci.lower = ci(time)["CI lower"], ci.upper = ci(time)["CI upper"])
ggplot(plot_data, aes(x = factor(textmethod), y = meanTime)) +
      geom_bar(stat="identity") +
      geom_errorbar(aes(ymin = ci.lower, ymax = ci.upper), width = .1, position = position_dodge(.05)) +
      theme(panel.background = element_blank(),
            panel.border = element_rect(colour = 'black', fill=NA),
            panel.grid.major.y = element_line( size = .1, color = "grey")
            ) +
      labs(title = "", x = "Text method", y = "Time")

```

## Pairwise comparisons

```{r posthoc}
attach(data.long)
pw <- pairwise.t.test(time, factor(textmethod), paired=F, p.adj = "bonferroni")
detach(data.long)
kable(pw$p.value)
```