---
title: "Repeated Measures ANOVA"
author: "Gery Casiez"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(apa)
```

## Repeated Measures ANOVA using 2 factors

Within-subject design 
Factor 1: text entry method (standard word-processing software, text entry using word-prediction software, text entry using speech-based dictation software)
Factor 2: task (transcription, composition)

## Load data

```{r readdata}
data = read.table("wordProcessingSoftwareWithinSubjectRepeatedMeasuresANOVA2factors.csv", header=TRUE, sep=",")
kable(data)
```


Convert data using the long format
```{r convertlong, results="hide", message=FALSE}
library("reshape")
data.long = melt(data, id = c("participant","textmethod","task","time"))
```

Indicate what are the independent variables (factors)
```{r indepvar, results="hide"}
data.long$textmethod = factor(data.long$textmethod)
data.long$task = factor(data.long$task)
```

## ANOVA
```{r anova}
library("ez")
kable(anova_apa(ezANOVA(data.long, dv=.(time), wid=.(participant), within=.(textmethod,task), detailed=TRUE)))
```

## Pairwise comparisons

```{r posthoc}
attach(data.long)
pw <- pairwise.t.test(time, factor(task), paired=T, p.adj = "bonferroni")
detach(data.long)
kable(pw$p.value)
```
